---
title: "My answers"
author: "My name"
date: 26 September 2018
output: pdf_document
---



Human papillomavirus (HPV) is the most common sexually transmitted infection in the United States. Some HPV types can cause genital warts and are considered low risk, with a small chance for causing cancer. Other types are considered high risk, causing cancer in different areas of the body
including the cervix and vagina in women, penis in men, and anus and oropharynx in both men and women. A report provides the most recent national estimates of genital HPV prevalence among adults aged 18–59 from the National Health and Nutrition Examination Survey (NHANES) 2013–2014. During 2013–2014, prevalence of any genital HPV was 42.5% among adults aged 18–59. This information applies to Questions 1-4.

#Question 1: Sampling distribution for a sample proportion

**1(a) If we use a sample proportion, $P$, based on a sample of size $n = 20$, to estimate the population proportion, $\pi = 0.425$, would it be very surprising to get an estimate that is off by more than 0.10 (that is, the sample proportion is less than 0.325 or greater than 0.525)? Support your answer.**


```{r}
# Write your answer here
```



**1(b) If we use a sample proportion, $P$, based on a sample of size $n = 100$, to estimate the population proportion, $\pi = 0.425$, would it be very surprising to get an estimate that is off by more than 0.10? Support your answer.**
```{r}
# Write your answer here
```



**1(d) If we use a sample proportion, $P$, based on a sample of size $n = 20$, to estimate the population proportion, $\pi = 0.425$, would it be very surprising to get an estimate that is off by more than 0.05 (that is, the sample proportion is less than 0.375 or greater than 0.475)? Support your answer.**
```{r}
# Write your answer here
```

**1(f) If we use a sample proportion, $P$, based on a sample of size $n = 500$, to estimate the population proportion, $\pi = 0.425$, would it be very surprising to get an estimate that is off by more than 0.05? Support your answer.**
```{r}
# Write your answer here
```




#Question 2: Maximum likelihood
Suppose we sample of size of $n=222$ random adult individuals age 18-59 and determine that 98 individuals were infected with genital HPV.  

**2(a) Draw the liklihood function for an estimate of the population parameter as a function of $\pi$ (with sample size $n=222$).**
```{r}
# Write your answer here
```

**2(c) Suppose that the sample size is increased to $n=444$ random adults between the age of 18-59 and observe that 196 individuals were infected with gential HPV. How does the likelihood function change from that in 2(a)? Explain and/or show.**
```{r}
# Write your answer here
```

#Question 3: Confidence interval for HPV
Suppose that you obtain a random sample of adult individuals age 18-59 and determine whether they were infected with genital HPV ($\pi = 0.425$). Use the `R` function `binom.test()`. 

**3(a) Use `R` to get a random sample of $n=20$ from adults  18-59. Determine the width of the exact 99\% confidence interval for $\pi$ for your sample. To get the width, you would subtract the lower bound of the 99\% confidence interval from the upper bound of the 99\% confidence interval.** 
```{r}
# Write your answer here
```


**3(c) Use `R` to get a random sample of $n=500$ from adults age 18-59. Determine the width of the exact 99\% confidence interval for $\pi$ for your sample.**
```{r}
# Write your answer here
```


#Question 4: Hypothesis test for HPV
Suppose that you obtain a random sample of $n = 123$ adult individuals age 18-59 and determine whether they were infected with genital HPV. Use the `R` function `binom.test()`. 

**4(a) Use `R` to get a random sample of $n=123$ from a population of adults age 18-59 with $\pi=0.405$. Report the sample proportion and perform a test to determine whether it differs from $\pi=0.425$** 
```{r}
# Write your answer here
```


**4(c)  Use `R` to get a random sample of $n=123$ from a population of adults age 18-59 with $\pi=0.385$. Report the sample proportion and perform a test to determine whether it differs from $\pi=0.425$**
```{r}
# Write your answer here
```



**4(e)  Use `R` to get a random sample of $n=123$ from a population of adults age 18-59 with $\pi=0.365$. Report the sample proportion and perform a test to determine whether it differs from $\pi=0.425$** 
```{r}
# Write your answer here
```



**4(g) Using parts information obtained in your answers above, comment on what you observe about the relationship between the sample proporiton, the hypothesized $\pi$, and the $p$-value.**
```{r}
# Write your answer here
```



**5(b) What is the probability that the death from smoking related disease will be between 238 and 245 cases in 1000 random women smokers? Again, compare the estimate obtained from a normal distribution to that obtained from the binomial distribution.**
```{r}
# Write your answer here
```



**5(d) What is the probability that the sample proportion of death from smoking related disease will be between 0.238 and 0.245 cases in 1000 random women smokers? Again, compare the estimate obtained from a normal distribution to that obtained from the binomial distribution.**
```{r}
# Write your answer here
```



**5(f) What is the probability that out of 24 women smokers between 5 and 9 will die from smoking related diseases? Use a normal approximation to the binomial to estimate this quantity and compare it to the exact probability from a binomial.**
```{r}
# Write your answer here
```



#Question 6: Confidence interval data analysis

Colon cancer is the most common gastrointestinal malignancy and the second-leading cause of cancer death in the United States. Approximately 80\% of colon cancer patients present with resectable, localized disease, and in these patients, nodal metastases have long been recognized as the most important factor predicting long-term survival. Nodal involvement is an important determinant in the decision to administer adjuvant chemotherapy, and with the demonstration over the last decade of highly effective systemic therapies for colon cancer, it is essential to ensure that all patients who would benefit from such treatment receive counseling concerning these therapies and have access to them. Numerous studies have shown an improvement in disease-specific and overall survival when increasing numbers of lymph nodes are examined for colon cancer. There has been a considerable effort to determine the minimum number of nodes that need to be evaluated to deem a patient free of nodal metastases with reasonable certainty. Estimates have varied from 6 to 40 lymph nodes; however, numerous studies and consensus guidelines have suggested that examination of 12 regional lymph nodes is a reasonable minimum for adequate nodal evaluation for colon cancer. Despite these findings, population-based assessments have shown that the majority of patients in the United States do not have 12 or more nodes examined. 

The American College of Surgeons (ACoS), American Society of Clinical Oncology (ASCO), and the National Comprehensive Cancer Network (NCCN) harmonized a quality measure requiring resection and pathological examination of 12 or more lymph nodes for colon cancer. Subsequently, the National Quality Forum (NQF) endorsed the 12-node measure for quality surveillance. A large, national health-care insurer, United Healthcare, is already basing referral recommendations for colectomy on a one-time requirement that surgeons provide pathology reports demonstrating examination of 12 or more lymph nodes for colon cancer. These organizations deem a hospital to be compliant with the 12-node measure if examination of at least 12 nodes occurred for at least 75\% of patients at that hospital. Due to statistical variation, hospitals were considered “statistically compliant” if the upper limit of the (two-sided) 95\% confidence interval (CI) of the estimate of their performance rate was greater than or equal to 75\%. 

**Dr. Smith has requested that you analyze the data provided to determine whether the hospital currently meets this cancer care quality metric for node examination in colorectal cancer patients.** 
Write a short paragraph that summarizes your findings and answers Dr. Smith's request.

*DATA*
The data are in a file called `crc.csv`

* `BMI`:	an indicator whether a person has a normal BMI (BMI less than or equal to 25) or an overweight BMI (BMI greater than 25)
* `Nodes`:	the number of lymph nodes examined
* `Year`:	year in which the patient underwent her/his colorectal cancer surgery

```{r}
# Write your answer here
```



**7(a) Is there evidence that the reoperation rate at Hopevale Hospital differs from that of their peer institutions? Use a significance level of 0.05.**
Perform an hypothesis test. Write your answer in a short paragraph that includes the p-value and an indication of how the rate differs (if it is statistically significant).
```{r}
# Write your answer here
```



**7(c) What is something that would be important to know regarding the tumor size at the peer hospitals?**
```{r}
# Write your answer here
```


